---
title: "Fixed- and random-effects meta-analysis in R"
subtitle: "A meta-analysis exercise - phytoestrogens and total cholesterol"
author: "Biostatistics, Epidemiology and Public Health - University of Padova"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: readable
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
# Global chunk options

knitr::opts_chunk$set(
  echo = TRUE, # show the code (this is a teaching document)
  message = FALSE, # hide package messages
  warning = FALSE, # hide package warnings
  fig.align = "center", # align figures
  collapse = TRUE # collapses code and output together in the final documen
)
```

# Research question

> Does phytoestrogen improve levels of total blood cholesterol in
> postmenopausal women?

| PICO         | In this exercise                              |
|--------------|-----------------------------------------------|
| Population   | Postmenopausal women                          |
| Intervention | Phytoestrogen                                 |
| Outcome      | Total blood cholesterol (mmol/L)              |
| Design       | Meta-analysis of randomized clinical trials   |

Each trial gives **one** estimate of the effect; the meta-analysis pools these
estimates into a single overall effect, giving more weight to the more precise
trials (larger samples, smaller variability).

# Effect measure

Each trial reports, in every arm, the **change** in total cholesterol
(follow-up − baseline).  
Because the outcome is continuous and on the same scale
(mmol/L), each trial is summarised by the **Mean Difference (MD)** between the
two changes:

$$\text{MD} = \text{Mean change}_{\text{intervention}} - \text{Mean change}_{\text{control}}
\qquad (\Delta = \text{follow-up} - \text{baseline})$$

Reading the sign of the MD:

* `MD < 0` → **larger** reduction of cholesterol in the phytoestrogen group
* `MD > 0` → **smaller** reduction (or an increase) in the phytoestrogen group
* `MD = 0` → **same average** change in both groups

# Packages

A package is a collection of ready-made functions.  
Install each one **once** per
computer, then load it at the start of every session.

```{r packages-install, eval=FALSE}
# Run ONCE, then leave commented out:
# install.packages(c("meta", "metafor", "tidyverse", "readxl"))
```

```{r packages-load}
library(meta)        # meta-analysis functions: metacont, forest, funnel, metabias, metainf
library(tidyverse)   # dplyr / readr / tibble: mutate, rename, near, the |> pipe
library(readxl)      # read_excel(): read the .xlsx data file
library(purrr)
```

The package `meta` contains all the functions used in this exercise, including:

* `metacont()` for fixed- and random-effects meta-analysis
* `forest()` for forest plots
* `funnel()` and metabias() for publication bias assessment
* `metainf()` for leave-one-out sensitivity analysis

# Load and explore the data

`read_excel()` (from **readxl**) loads the "MA exercise" workbook into a
*tibble* - a modern data frame.  
We read the sheet `"Total cholesterol"`, which contains the data used for the meta-analysis.  
Each row corresponds to one randomized clinical trial.  
The second sheet contains the data dictionary, describing the variables included in the dataset.

```{r load-data}
dat <- read_excel("data/MA_exercise.xlsx",
                  sheet = "Total cholesterol")  # adjust the path if needed

glimpse(dat)            # inspect variable names, types, and example values
```

`glimpse()` shows the structure of the dataset:
- `<chr>` indicates character variables (text),
- `<dbl>` indicates numeric variables

Always check that variables have the expected type before starting the analysis.

```{r}
slice_head(dat, n = 6)  # display the first 6 rows of the dataset
```

# Understand the dataset

The dataset contains 25 randomized clinical trials and 42 variables.
For the meta-analysis we only need the study label, sample sizes, mean changes,
and standard deviations. The remaining variables describe study characteristics
(e.g. age, BMI, intervention type, duration, location, study quality) and may
be useful later to explore heterogeneity.  
Each row is one trial.  
The study label arrives as `Author, Year of Publication`;
we rename it to `study` in the next step.  
For each arm the file stores the value
at **baseline**, at **follow-up**, and the **change** between them:

| Column(s) | Meaning |
|-----------|---------|
| `n_int`, `n_control` | Sample size of the intervention and control groups |
| `int_base_mean`, `int_base_sd` | Mean and SD of total cholesterol at baseline, intervention group |
| `int_final_mean`, `int_final_sd` | Mean and SD of total cholesterol at follow-up, intervention group |
| `int_mean_diff`, `int_sd_diff` | Mean and SD of the change (follow-up − baseline), intervention group |
| `cont_base_mean`, `cont_base_sd` | Mean and SD of total cholesterol at baseline, control group |
| `cont_final_mean`, `cont_final_sd` | Mean and SD of total cholesterol at follow-up, control group |
| `cont_mean_diff`, `cont_sd_diff` | Mean and SD of the change (follow-up − baseline), control group |

The columns that actually enter the analysis are the **change** ones
(`*_mean_diff`, `*_sd_diff`) plus the group sizes - the baseline/follow-up
columns are kept so we can see where the change comes from.

# Compute the study-specific effect size

**Design choice (compute vs. pre-compute the MD).**  
The file keeps the raw per-arm summary statistics because `metacont()` needs them to weight each trial
by its precision.  
The change scores `int_mean_diff` / `cont_mean_diff` are
themselves pre/post differences (follow-up − baseline), so we can recompute them
from the baseline/final columns and **check** they match.  
We then form the
study-level MD by hand, only to *see* the effect size and its sign - this `MD`
column is **not** passed to `metacont()` (the function recomputes it internally).

```{r compute-md}
dat <- dat |>
  rename(study = `Author, Year of Publication`) |>     # tidy the study-label column
  mutate(
    int_change  = int_final_mean  - int_base_mean,      # recompute the change, intervention arm
    cont_change = cont_final_mean - cont_base_mean,      # recompute the change, control arm
    MD          = int_mean_diff   - cont_mean_diff       # study effect size (understanding only)
  )

# Check our recomputed change matches the change score already in the file.
# near() is a tolerant equality test (avoids a false FALSE from rounding).
test <- dat |>
  mutate(ok_int  = near(int_change,  int_mean_diff),
         ok_cont = near(cont_change, cont_mean_diff)) |>
  count(ok_int, ok_cont)                                 # ideally all TRUE

# # Display the studies and visually inspect their calculated mean differences
test1 <- dat |>
  select(study, int_mean_diff, cont_mean_diff, MD) |>   # the building blocks of the MD
  slice_head(n = 6)
```

Worked example - **Aubertin et al, 2007**:  
change in the treated group $= -0.14$,  
change in the control group $= -0.09$,  
so $\text{MD} = -0.14 - (-0.09) = -0.05$ mmol/L.  
The negative MD means the phytoestrogen group had a **larger reduction** in total cholesterol than the
control group.

> **Note.** Only the *means* can be recomputed as `follow-up − baseline`. The SD
> of the change (`int_sd_diff` / `cont_sd_diff`) **cannot** be derived from the
> baseline/final SDs without the pre/post correlation, so we use it as given.
> The optional section below explains why.

## Preparing the moderators (data cleaning)

The covariate columns are typed by hand, so the labels are messy
(`China` vs `China `, `usa` vs `USA`, a dozen spellings of *isoflavone*…).
We tidy them once with **dplyr** before grouping — cleaning is part of the job.

```{r prep-moderators}
dat <- dat |>
 mutate(
    quality = factor(str_trim(`study quality`), levels = c("poor","fair","good")),
    duration_wk  = duration12,
    duration_grp = factor(if_else(duration12 > 12, "> 12 weeks", "<= 12 weeks")),
    region  = recode(str_to_title(str_trim(location)), "Usa"="USA", "Uk"="UK"),
    agegrp  = factor(Stratum1, labels = c("<=55 y", ">55 y")),
    menop   = factor(Stratum3, labels = c("<= median", "> median")),
    bmigrp  = factor(Stratum2, labels = c("<= median", "> median")),
    admin   = factor(pills1dietary0, labels = c("Dietary", "Pills")),
    health  = factor(healthy1, labels = c("With condition", "Healthy")),
    smoking = factor(nonsmok1, labels = c("Smokers/mixed", "Non-smokers")),
    ptype = factor(case_when(
      str_detect(str_to_lower(type), "lignan|flax|sesame") ~ "Lignan",
      str_detect(str_to_lower(type), "clover|coumestan")    ~ "Coumestan",
      str_detect(str_to_lower(type), "iso|soy|genistein")   ~ "Isoflavone/soy",
      TRUE                                                  ~ "Other"),
      levels = c("Isoflavone/soy","Lignan","Coumestan","Other")),
    baseline_tc = (int_base_mean*n_int + cont_base_mean*n_control) / (n_int + n_control)
  )
```


```{r}
dat <- dat |>
  group_by(study) |>
  mutate(study = if (n() > 1) paste0(study, " (", row_number(), ")") else study) |>
  ungroup()
```

# Fixed-effects model

`metacont()` is the meta-analysis of **cont**inuous outcomes.  
It takes the per-arm summary statistics, computes each trial's MD and precision, and pools
them.  
The **fixed (common) effect** model assumes *one* single true effect shared by every trial; trials differ only by random sampling error, so larger /
more precise trials get more weight.

```{r fixed-model}
# Fit a fixed-effect meta-analysis using the Mean Difference (MD)
m_fixed <- metacont(
  n.e     = n_int,           # n.e    = size of the Experimental (intervention) arm
  mean.e  = int_mean_diff,   # mean.e = mean change, experimental (intervention) arm
  sd.e    = int_sd_diff,     # sd.e   = SD of the change, experimental (intervention) arm
  n.c     = n_control,       # n.c    = size of the Control arm
  mean.c  = cont_mean_diff,  # mean.c = mean change, control arm
  sd.c    = cont_sd_diff,    # sd.c   = SD of the change, control arm
  studlab = study,           # studlab = label shown for each study
  data    = dat,             # data    = the tibble holding the columns above
  sm      = "MD",            # sm      = Summary Measure -> raw Mean Difference
  # sm = "SMD"   # Standardized Mean Difference
  # sm = "ROM"   # Ratio of Means
  common  = TRUE,            # common  = TRUE  -> fit the FIXED (common) effect model
  random  = FALSE,            # random  = FALSE -> skip the random-effects model here
  method.tau = "DL"          
)

summary(m_fixed)             # pooled effect, 95% CI and heterogeneity statistics
```

**What is the pooled mean difference?**  
The fixed-effect pooled mean difference is
**`r sprintf("%.2f", m_fixed$TE.common)` mmol/L**.

**Is the result statistically significant?**  
Yes.  
The 95% confidence interval ranges from
**`r sprintf("%.2f", m_fixed$lower.common)`** to
**`r sprintf("%.2f", m_fixed$upper.common)`**, and does not include 0.

**What does the confidence interval tell us?**  
The confidence interval is very narrow, indicating a precise estimate of the
overall effect. Because the entire interval is below 0, the phytoestrogen group
experienced a significantly greater reduction in total cholesterol than the
control group.

Under the fixed-effect model, larger and more precise studies receive more
weight, so a single large study can have a substantial influence on the pooled
estimate.

## Forest plot (fixed-effect model)

A forest plot is the standard picture of a meta-analysis: each trial is a square
(its MD) with a horizontal line (its 95% CI); the square's size shows the
trial's weight, and the diamond at the bottom is the pooled effect.

```{r forest-fixed, fig.width=9, fig.height=11}
# Forest plot:
# - each square represents a study-specific effect estimate (MD)
# - larger squares indicate studies with greater weight
# - horizontal lines show the 95% confidence intervals
# - the diamond represents the pooled effect estimate

forest(m_fixed,
       common      = TRUE,                  # draw the fixed-effect diamond
       label.left  = "Favours treatment",   # text under the LEFT end of the axis
       label.right = "Favours control",     # text under the RIGHT end of the axis
       xlab        = "Mean difference in total cholesterol (mmol/L)",
       leftcols    = c("studlab", "n.e", "n.c"),
       leftlabs    = c("Study", "N (treat)", "N (ctrl)"))
```

**How to read the forest plot**

- Each square represents the effect estimate from a single study.
- Larger squares indicate studies that contribute more weight to the pooled estimate.
- The horizontal line through each square is the 95% confidence interval.
- The diamond at the bottom represents the pooled effect; its width corresponds to the 95% confidence interval.
- The solid vertical line represents the line of no effect (MD = 0).
- The dashed vertical line represents the pooled effect estimate from the meta-analysis.

**Interpretation**

Women receiving phytoestrogen supplementation experienced a greater reduction in
total cholesterol than women in the control group.

Under the fixed-effect model, we assume that all studies estimate the same true
underlying effect and that observed differences between studies are due only to
random sampling error.

# Random-effects model

The fixed-effect model assumes a single common effect.  
We now relax this assumption and allow the true effect to vary across studies. 
Same call, but we switch the model. The **random-effects** model assumes the
true effect is *not* identical across trials but varies around an average
(doses, types of phytoestrogen and populations differ). It adds the
between-study variance $\tau^2$ to each weight, which usually **widens** the
confidence interval.

```{r random-model}
# Fit a random-effects meta-analysis using the Mean Difference (MD)
m_random <- metacont(
  n.e     = n_int,
  mean.e  = int_mean_diff,
  sd.e    = int_sd_diff,
  n.c     = n_control,
  mean.c  = cont_mean_diff,
  sd.c    = cont_sd_diff,
  studlab = study,
  data    = dat,
  sm      = "MD",
  common     = FALSE,        # common = FALSE -> skip the fixed-effect model
  random     = TRUE,         # random = TRUE  -> fit the RANDOM-effects model
  method.tau = "DL"          # method.tau = "DL" -> DerSimonian-Laird estimator of tau^2
                             # (the default in 'meta' is REML; we use DL to match the exercise)
                             # Only relevant for random-effects models, since fixed-effect models assume τ² = 0
)

summary(m_random) # pooled estimate, 95% CI, and heterogeneity statistics
```

**What is the pooled mean difference?**  
The random-effects pooled mean difference is
**`r sprintf("%.2f", m_random$TE.random)` mmol/L**.

**Is the result statistically significant?**  
Yes. The 95% confidence interval ranges from
**`r sprintf("%.2f", m_random$lower.random)`** to
**`r sprintf("%.2f", m_random$upper.random)`**, and does not include 0.

**What does the confidence interval tell us?**  
The confidence interval is wider than in the fixed-effect model because the
random-effects model incorporates both within-study variability and
between-study variability (heterogeneity). The entire interval remains below 0,
indicating that the phytoestrogen group experienced a significantly greater
reduction in total cholesterol than the control group.

Unlike the fixed-effect model, the random-effects model assumes that the true
effect may vary across studies and therefore provides an estimate of the average
effect across a distribution of study-specific effects.

**Compare the random-effects model with the fixed-effect model**

**Is the pooled estimate different?**  
Yes. The pooled estimate from the random-effects model is
**`r sprintf("%.2f", m_random$TE.random)` mmol/L**, compared with
**`r sprintf("%.2f", m_fixed$TE.common)` mmol/L** under the fixed-effect model.

**Is the confidence interval wider or narrower?**  
The confidence interval is wider under the random-effects model:
**`r sprintf("%.2f", m_random$lower.random)` to `r sprintf("%.2f", m_random$upper.random)`**
versus
**`r sprintf("%.2f", m_fixed$lower.common)` to `r sprintf("%.2f", m_fixed$upper.common)`**
for the fixed-effect model.

**Why?**  
The fixed-effect model assumes that all studies estimate the same true effect and
accounts only for within-study variability. The random-effects model allows the
true effect to vary across studies and incorporates both within-study and
between-study variability (heterogeneity). As a result, uncertainty increases
and the confidence interval becomes wider.

## Forest plot (random-effect model)

```{r forest-random, fig.width=9, fig.height=11}
# Forest plot:
# - each square represents a study-specific effect estimate (MD)
# - larger squares indicate studies with greater weight
# - horizontal lines show the 95% confidence intervals
# - the diamond represents the pooled effect estimate
forest(m_random,
       random      = TRUE,                  # draw the random-effects diamond
       label.left  = "Favours treatment",
       label.right = "Favours control",
       xlab        = "Mean difference in total cholesterol (mmol/L)",
       leftcols    = c("studlab", "n.e", "n.c"),
       leftlabs    = c("Study", "N (treat)", "N (ctrl)"))
```

**Interpretation**

The random-effects model also suggests a beneficial effect of phytoestrogens on
total cholesterol.

However, unlike the fixed-effect model, it assumes that the true effect may vary
across studies because populations, interventions and study settings differ. The
pooled estimate therefore represents the average effect across a distribution of
true effects.

# Assess heterogeneity

Heterogeneity measures how much the true effect varies between trials:

* **Q** - Cochran's Q (total variation); its *p*-value tests whether the
  differences exceed chance.
* $\tau^2$ - estimated between-study variance (0 = the trials all agree).
* $I^2$ - % of total variation that is real between-study difference rather than
  chance. Rough guide: 0–25% low, 25–50% moderate, 50–75% substantial,
  >75% considerable.

```{r heterogeneity}
# Display the main heterogeneity statistics:
# - I²: percentage of variability due to heterogeneity
# - τ²: between-study variance
# - Q: Cochran's heterogeneity statistic
# - p-value: test of whether heterogeneity exceeds chance

cat("I-squared:", m_random$I2, "%\n")

cat("Tau-squared:", round(m_random$tau2, 4), "\n") # τ² = estimated between-study variance

cat("Cochran's Q:", round(m_random$Q, 1), "\n") # Q = Cochran's heterogeneity statistic

cat("P-value for Q:", signif(m_random$pval.Q, 3), "\n")
# p-value for the Q test; a small p-value suggests heterogeneity beyond chance
```

Here $I^2 \approx$ **`r sprintf("%.1f", m_random$I2 * 100)`%** - considerable
heterogeneity. The trials genuinely differ, so the fixed-effect assumption of
one shared effect is not credible.

# Compare the two models

| Fixed-effect                          | Random-effects                          |
|---------------------------------------|-----------------------------------------|
| Assumes one true effect               | Assumes different true effects          |
| More weight to large studies          | More balanced weighting                 |
| Narrower confidence interval          | Wider confidence interval               |
| Appropriate when heterogeneity is low | Preferred when heterogeneity is present |

For this analysis the **random-effects** model is more appropriate: the trials
used different inclusion/exclusion criteria and different types/doses of
phytoestrogen, and the high heterogeneity
($I^2 \approx$ `r sprintf("%.1f", m_random$I2 * 100)`%) confirms the effect is
not shared.


# Publication bias

Small studies with no effect are sometimes never published, biasing the pooled
result. A funnel plot shows each trial's effect against its precision; without
bias the points form a symmetric inverted funnel. `metabias()` tests that
symmetry with Egger's regression test.

```{r publication-bias, fig.width=7, fig.height=6}
# Funnel plot: study effect size versus study precision
# In the absence of publication bias, points should form a symmetric funnel
funnel(m_random, studlab = FALSE)            # funnel plot (no per-point labels)
```

```{r}
# Egger's regression test for funnel plot asymmetry
# A small p-value suggests evidence of asymmetry
metabias(m_random, method.bias = "Egger")    # Egger's test for funnel asymmetry
```


**How to interpret the results**

- In the funnel plot, a symmetric inverted funnel suggests little evidence of publication bias.
- Asymmetry may indicate publication bias, although other factors (e.g. heterogeneity or small-study effects) can also produce asymmetry.
- Egger's test formally assesses funnel plot asymmetry.
- A small p-value (commonly < 0.05) suggests evidence of asymmetry.

Publication bias assessments should be interpreted cautiously, especially when
the number of studies is small.

# (Optional) Where do the change-score SDs come from?

A paper rarely reports the SD of the change directly. Depending on what is
reported, it is reconstructed in one of three ways. This is background for the
curious - for the exercise itself we just use the `*_sd_diff` columns as given.

```{r sd-methods}
# METHOD 1 - Reported directly: the trial publishes the change and its SD.
#            Nothing to compute; use the column as given.
dat |> 
  select(study, int_mean_diff, int_sd_diff) |> 
  slice_head(n = 6)

# METHOD 2 - Derived from the precision of the change (SE or 95% CI):
#   from SE:     SD = SE * sqrt(n)
#   from 95% CI: SD = sqrt(n) * (upper - lower) / (2 * 1.96)
#   (illustrative only - these SE/CI columns are NOT in this dataset)
# dat |>
#   mutate(sd_from_se = int_change_se * sqrt(n_int),
#          sd_from_ci = sqrt(n_int) * (int_change_ci_upp - int_change_ci_low) / (2 * 1.96))

# METHOD 3 - Imputed from baseline & final SDs (needs a pre/post correlation r):
#   SD_change = sqrt(SD_base^2 + SD_final^2 - 2*r*SD_base*SD_final)
#   r is NOT in the data, so it is ASSUMED (Cochrane Handbook default: r = 0.5).
r <- 0.5
dat |>
  mutate(sd_change_imp = sqrt(int_base_sd^2 + int_final_sd^2 -
                              2 * r * int_base_sd * int_final_sd)) |>
  select(study, int_sd_diff, sd_change_imp) |> slice_head(n = 6)
```

**Reverse check - what correlation is implied by this file?** Invert Method 3 to
solve for `r`. A value outside `[-1, 1]` means the three SDs are *not* internally
consistent with a standard pre/post calculation.

```{r sd-reverse-check}
dat |>
  mutate(
    r_int  = (int_base_sd^2  + int_final_sd^2  - int_sd_diff^2)  /
             (2 * int_base_sd  * int_final_sd),
    r_cont = (cont_base_sd^2 + cont_final_sd^2 - cont_sd_diff^2) /
             (2 * cont_base_sd * cont_final_sd)
  ) |>
  select(study, r_int, r_cont) |> slice_head(n = 10)
```

In this dataset the implied correlations come out around 1 (some even above 1,
which is impossible). So the change-score SDs are **not** internally consistent
with the baseline/final SDs - they were obtained in another way, and we use them
as provided rather than trying to reconstruct them. (A value above 1 is the data
telling us something, not a mistake on our side.)



# Take-home messages

1. Fixed- and random-effects models rest on **different assumptions** - knowing
   them tells you which fits your research question.
2. They are interpreted differently: one true effect for all (fixed) vs. an
   average of effects that vary across studies (random).
3. Always inspect heterogeneity ($I^2$, $Q$, $\tau^2$): it guides **both** the
   choice of model and the interpretation.

**Answers to the exercise**

* Fixed-effects result: MD = `r sprintf("%.2f", m_fixed$TE.common)` mmol/L
  (95% CI `r sprintf("%.2f", m_fixed$lower.common)` to
  `r sprintf("%.2f", m_fixed$upper.common)`).
* Random-effects result: MD = `r sprintf("%.2f", m_random$TE.random)` mmol/L
  (95% CI `r sprintf("%.2f", m_random$lower.random)` to
  `r sprintf("%.2f", m_random$upper.random)`).
* Same direction (cholesterol is lowered) but the random interval is wider.
* Appropriate model: **random effects**, because heterogeneity is very high
  ($I^2 \approx$ `r sprintf("%.1f", m_random$I2 * 100)`%).

---

# Investigating heterogeneity

This section prepares the analyses presented in the follow-up
lecture (*Dealing with Heterogeneity*) on the **same** dataset.
These analyses are run on the random-effects model
because that is the appropriate one here (I² ≈ 97.5%).


## Subgroup analysis

A subgroup analysis splits the trials into groups and asks whether the pooled
effect **differs across groups**. The right tool is the **test for subgroup
differences** (a between-group Q test), *not* eyeballing whether the
subgroup CIs overlap.

```{r subgroup-type}
# Subgroup by age
m_sub_age <- metacont(
  n.e = n_int, 
  mean.e = int_mean_diff, 
  sd.e = int_sd_diff,
  n.c = n_control, 
  mean.c = cont_mean_diff, 
  sd.c = cont_sd_diff,
  studlab = study, 
  data = dat, sm = "MD",
  common = FALSE, 
  random = TRUE, 
  method.tau = "DL",
  subgroup = agegrp, # subgroup = the grouping variable (a factor)
  subgroup.name = "Median age, y" # subgroup.name = label shown in the output
)


m_sub_age        # read: per-subgroup MD, and the "Test for subgroup differences"
```

```{r}
# Helper that runs ONE subgroup analysis for a given variable name `v`
# and returns its results as a tidy one-row-per-subgroup table.
run_sub <- function(v) {
  m <- metacont(
    n.e = n_int, mean.e = int_mean_diff, sd.e = int_sd_diff,
    n.c = n_control, mean.c = cont_mean_diff, sd.c = cont_sd_diff,
    studlab = study, data = dat, sm = "MD",
    common = FALSE, random = TRUE, method.tau = "DL",
    subgroup = .data[[v]],      # .data[[v]] = use the column whose name is in v
    subgroup.name = v
  )
  # Pull the per-subgroup results out of the fitted object (the ".w" = within-subgroup)
  tibble(
    variable  = v,
    subgroup  = m$bylevs,                       # the subgroup levels
    k         = m$k.w,                          # number of studies in each subgroup
    MD        = round(m$TE.random.w, 3),        # pooled MD within the subgroup
    lower     = round(m$lower.random.w, 3),     # 95% CI lower bound
    upper     = round(m$upper.random.w, 3),     # 95% CI upper bound
    I2        = round(m$I2.w * 100, 1),         # residual heterogeneity within the subgroup
    p_between = round(m$pval.Q.b.random, 3)     # test for subgroup differences (the key number)
  )
}
```

```{r}
# Variables we want to test as subgroups (all are factors created above)
group_vars <- c("ptype","duration_grp","quality","region",
                "agegrp","menop","bmigrp","admin","health","smoking")

# Run run_sub() on each, stacking the results into one table.
# possibly() = if one variable fails (e.g. a single level), skip it instead of stopping.
subgroup_table <- map_dfr(group_vars, possibly(run_sub, otherwise = NULL))
print(subgroup_table, n = 100)   # the column that matters is p_between
```


```{r subgroup-forest, fig.width=9, fig.height=12}
forest(m_sub_age)   # forest plot split by subgroup, with subgroup diamonds
```
## Meta-regression

Meta-regression extends subgroup analysis to **continuous** moderators: it
regresses the study effect on a study-level characteristic and asks whether that
characteristic explains part of the heterogeneity. (Rule of thumb: only with
≥ ~10 studies.)

```{r metareg-baseline}
# Does the effect depend on the trial's BASELINE cholesterol?
# =====================================================================
#  META-REGRESSION
#  Does a continuous study-level characteristic explain the heterogeneity?
#  (Rule of thumb: only with ~10+ studies.)
# =====================================================================

# --- (a) Baseline cholesterol -- the informative moderator -----------
# Trials whose participants start higher: do they show a larger reduction?
mr_base <- metareg(m_random, ~ baseline_tc)
mr_base          # look at the slope, its p-value, and R^2 (heterogeneity explained)
```

```{r}
# Bubble plot: each bubble = a study (size = precision), line = regression fit
bubble(mr_base, studlab = FALSE,
       xlab = "Baseline total cholesterol (mmol/L)",
       ylab = "Mean difference (mmol/L)")

```

```{r}
# --- (b) Duration -- a "clean" (non-coupled) moderator ---------------
# Not coupled to the effect size, so no regression-to-the-mean artefact.
mr_dur <- metareg(m_random, ~ duration_wk)
mr_dur
```

```{r}
# Bubble plot: each bubble = a study (size = precision), line = regression fit
bubble(mr_dur, studlab = FALSE,
       xlab = "Duration week",
       ylab = "Mean difference (mmol/L)")
```



```{r}
# --- (c) Age -- another non-coupled moderator ------------------------
mr_age <- metareg(m_random, ~ Age)
mr_age
```

```{r}
# Bubble plot: each bubble = a study (size = precision), line = regression fit
bubble(mr_age, studlab = FALSE,
       xlab = "Age, year",
       ylab = "Mean difference (mmol/L)")
```


**Interpretation** On these 25 trials, the only moderator with a signal is
**baseline cholesterol**: trials starting from higher cholesterol tend to show a
larger reduction (slope ≈ **−0.09** mmol/L per mmol/L, **p ≈ 0.11**, **R² ≈ 12%**).
**Duration** and **age** do *not* explain the heterogeneity (slope ≈ 0, large p,
R² ≈ 0%) — consistent with the paper, which found no duration effect in
meta-regression despite the subgroup signal. Two important caveats: baseline
cholesterol is **mathematically coupled** to the change (the effect is
follow-up − baseline, so using baseline as predictor can induce *regression to
the mean*); and meta-regression is **observational at the study level** — an
association between study characteristics and effect, **not** proof of causation.




## Sensitivity analysis

A sensitivity analysis repeats the meta-analysis under **alternative reasonable
choices**, to check the result is not driven by one decision or one study.

```{r sens-leave-one-out, fig.width=9, fig.height=12}
# 1) Leave-one-out: re-pool dropping one trial at a time
# Sensitivity 1) Leave-one-out: re-pool the random-effects model 25 times,
# each time dropping ONE trial, to see whether any single study drives the result.
inf <- metainf(m_random, pooled = "random")

inf            # numeric table: pooled MD with each study removed

forest(inf)    # visual: if all rows line up, no single study is influential
```

```{r sens-exclude-quality}
# Sensitivity 2) Restrict to higher-quality trials:
# subset = keep only studies whose quality is not "poor", then re-pool.
m_no_poor <- update(m_random, subset = (quality != "poor"))

m_no_poor      # compare this pooled MD with the main one
```

```{r sens-estimator}
# Sensitivity 3) Change the heterogeneity estimator (an "arbitrary" choice):
# refit with REML instead of DerSimonian-Laird and compare side by side.
m_reml <- update(m_random, method.tau = "REML")
c(DL_MD   = m_random$TE.random, DL_low   = m_random$lower.random, DL_upp   = m_random$upper.random,
  REML_MD = m_reml$TE.random,   REML_low = m_reml$lower.random,   REML_upp = m_reml$upper.random)
```

**How to read it:** if the pooled estimate and its significance stay essentially
the same across all three checks (no single study flips it, excluding poor-quality
trials doesn't change the message, DL ≈ REML), the conclusion is **robust** and
can be reported with greater confidence.


```{r session-info}
sessionInfo()   # records package versions, so the report is reproducible
```
